Subsetting
===========

Subsetting - Basics
----------

* [ always returns an object of the same class as the original; can be used to select more than one element (there is one exception)
* [[ is used to extract elements of a list of data frame; it can only be used to extract a single element and the class of the returned object will not necessarily be a list or data frame
* $ is used to extract elements of a list or data frame by name; semantics are similar to that of [[

```{r}
x <- c("a", "b", "c", "c", "d", "a")
x[1]
x[1:4]
x[x > "a"]
u <- x > "a"
x[u]
```

Subsetting - Lists
-------------------

```{r}
x <- list(foo = 1:4, bar = 0.6)
x[1]
x[[1]]
x$bar
x[["bar"]]
x["bar"]

x <- list(foo = 1:4, bar = 0.6, baz = "hello")
x[c(1, 3)]
```

The [[ operator can be used with computed indices; $ can only be used with literal names.

```{r}
x <- list(foo = 1:4, bar = 0.6, baz = "hello")

name <- "foo"
x[[name]]
x$name
x$foo
```

Subsetting Nested Elements of a List
------------------------------------

The [[ can take an integer sequence.

```{r}
x <- list(a = list(10, 12, 14), b = c(3.14, 2.81))
x[[c(1, 3)]]
# is the same as
x[[1]][[3]]
x[[c(2, 1)]]
```

Subsetting a Matrix
-------------------

Matrices can be subsetted in the usual way with (i,j) type indices.

```{r}
x <- matrix(1:6, 2, 3)
x[1, 2]
x[2, 1]
```

Indices can also be missing.

```{r}
x <- matrix(1:60, 12, 5)
x[1,]
x[,2]
```

By default when a single element is retrieved, it is returned as a vector of length 1 rather than a 1 x 1 matrix. This behavior can be turned off by setting drop = FALSE.

```{r}
x <- matrix(1:6, 2, 3)
x[1, 2]
x[1, 2, drop = FALSE]
```

Similarly, subsetting a single column or a single row will give you a vector, not a matrix (by default)

```{r}
X <- matrix(1:6, 2, 3)
x[1,]
x[1,,drop = FALSE]
```

Partial Matching
----------------

Partial matching of names is allowed with [[ and $.
```{r}
x <- list(phenomenal = 1:5)
x$p
x[["p"]]
x[["p", exact = FALSE]]
```

Removing NA Values
------------------

A common task is to remove missing values (NAs).
```{r}
x <- c(1, 2, NA, 4, NA, 5)
bad <- is.na(x)
x[!bad]
```

What if there are multiple things and you want to take the subset with no missing values?
```{r}
y <- c("a", "b", NA, "d", NaN, "f")
good <- complete.cases(x, y); good
x[good]
y[good]
```