---
title: "Week 1 Quiz"
output: statsr:::statswithr_lab
---

## Question 1
1. R was developed by statisticians working at:
<ol>
<li> **Answer:** The University of Auckland </li>
<li> **Explanation:** The R language was developed by Ross Ihaka and Robert Gentleman who were statisticians at the University of Auckland in New Zealand. </li>
</ol>

## Question 2
2. The definition of free software consists of four freedoms (freedoms 0 through 3). Which of the following is NOT one of the freedoms that are part of the definition? Select all that apply.
<ol>
<li> **Answer 1:** The freedom to prevent users from using the software for undesirable purposes.  </li>
<li> **Explanation:** This is not part of the free software definition. Freedom 0 requires that the users of free software be free to use the software for any purpose. </li>
<li> . </li>
<li> **Answer 2:** The freedom to sell the software for any price. </li>
<li> **Explanation:** This is not part of the free software definition. The free software definition does not mention anything about selling software (although it does not disallow it). </li>
<li> . </li>
<li> **Answer 3:** The freedom to restrict access to the source code for the software. </li>
<li> **Explanation:** This is not part of the free software definition. Freedoms 1 and 3 require access to the source code. </li>
</ol>

## Question 3
3. In R the following are all atomic data types EXCEPT: (Select all that apply)
<ol>
<li> **Answer:** list, array, data frame, matrix, and table </li>
<li> **Explanation:** As said in the lecture on data types, there are the main atomic data types: logical, complex, character, integer, and numeric </li>
</ol>

## Question 4
4. If I execute the expression x <- 4 in R, what is the class of the object `x' as determined by the `class()' function?
<ol>
<li> **Answer:** numeric </li>
<li> **Explanation:**
```{r}
x <- 4; class(x)
```
</li>
</ol>

# Question 5
5. What is the class of the object defined by x <- c(4, TRUE)?
<ol>
<li> **Answer:** numeric </li>
<li> **Explanation:**
```{r}
x <- c(4, TRUE); class(x)
```
</li>
</ol>

# Question 6
6. If I have two vectors x <- c(1,3, 5) and y <- c(3, 2, 10), what is produced by the expression cbind(x, y)?
<ol>
<li>**Answer:** matrix of 2 col and 3 row</li>
<li>**Explanation:**
```{r}
x <- c(1, 3, 5); y <- c(3, 2, 10)
cbind(x, y)
```
</li>
</ol>

# Question 7
7. A key property of vectors in R is that
<ol>
<li> **Answer:** elements of a vector all must be of the same class </li>
<li> **Explanation:** The most basic object is a vector: can only contain lists of the same class. (Taken from the lecture slide) </li>
</ol>

# Question 8
8. Suppose I have a list defined as x <- list(2, "a", "b", TRUE). What does x[[1]] give me? Select all that apply.
<ol>
<li> **Answer:** a numeric vector containing the element 2 || a numeric vector of length 1 </li>
<li> **Explanation:** 
```{r}
x <- list(2, "a", "b", TRUE); x[[1]]
```
</li>
</ol>

# Question 9
9. Suppose I have a vector x <- 1:4 and y <- 2:3. What is produced by the expression x + y?
<ol>
<li> **Answer:** an integer vector with the values 3, 5, 5, 7 </li>
<li> **Explanation:** 
```{r}
x <- 1:4; y <- 2:3
x + y
```
</li>
</ol>

# Question 10
10. Suppose I have a vector x <- c(17, 14, 4, 5, 13, 12, 10) and I want to set all elements of this vector that are greater than 10 to be equal to 4. What R code achieves this? Select all that apply.
<ol>
<li> **Answer:** x[x > 10] <- 4 || x[x >= 11] <- 4</li>
<li> **Explanation:**
```{r}
x <- c(17, 14, 4, 5, 13, 12, 10)
x[x > 10] <- 4
x[x >= 11] <- 4
```
</li>
</ol>

# Question 11
11. In the dataset provided for this Quiz, what are the column names of the dataset?
<ol>
<li> **Answer:** "Ozone"   "Solar.R" "Wind"    "Temp"    "Month"   "Day"</li>
<li> **Explanation:**
```{r}
data <- read.csv("hw1_data.csv")
names(data)
```
</li>
</ol>

# Question 12
12. Extract the first 2 rows of the data frame and print them to the console. What does the output look like?
<ol>
<li> **Answer:**
```{r echo=FALSE}
data <- read.csv("hw1_data.csv")
head(data, 2)
```
</li>
<li> **Explanation:**
```{r}
data <- read.csv("hw1_data.csv")
head(data, 2)
```
</li>
</ol>

# Question 13
13. How many observations (i.e. rows) are in this data frame?
<ol>
<li> **Answer:** 153 </li>
<li> **Explanation:**
```{r}
data <- read.csv("hw1_data.csv")
nrow(data)
```
</li>
</ol>

# Question 14
14. Extract the last 2 rows of the data frame and print them to the console. What does the output look like?
<ol>
<li> **Answer:** 
```{r echo=FALSE}
data <- read.csv("hw1_data.csv")
tail(data, 2)
```
 </li>
<li> **Explanation:**
```{r}
data <- read.csv("hw1_data.csv")
tail(data, 2)
```
</li>
</ol>

# Question 15
15. What is the value of Ozone in the 47th row?
<ol>
<li> **Answer:** 21 </li>
<li> **Explanation:**
```{r}
data <- read.csv("hw1_data.csv")
data[47,]
```
</li>
</ol>

# Question 16
16. How many missing values are in the Ozone column of this data frame? 
<ol>
<li> **Answer:** 37 </li>
<li> **Explanation:**
```{r}
data <- read.csv("hw1_data.csv")
nrow(subset(data, is.na(Ozone)))
```
</li>
</ol>

# Question 17
17. What is the mean of the Ozone column in this dataset? Exclude missing values (coded as NA) from this calculation.
<ol>
<li> **Answer:** 42.1 </li>
<li> **Explanation:**
```{r}
data <- read.csv("hw1_data.csv")
subs <- subset(data, !is.na(Ozone), select = Ozone)
apply(subs, 2, mean)
```
</li>
</ol>

# Question 18
18. Extract the subset of rows of the data frame where Ozone values are above 31 and Temp values are above 90. What is the mean of Solar.R in this subset? 
<ol>
<li> **Answer:** 212.8 </li>
<li> **Explanation:**
```{r}
data <- read.csv("hw1_data.csv")
subs <- subset(data, Ozone > 31 & Temp > 90, select = Solar.R)
apply(subs, 2, mean)
```
</li>
</ol>

# Question 19
19. What is the mean of "Temp" when "Month" is equal to 6?
<ol>
<li> **Answer:** 79.1 </li>
<li> **Explanation:**
```{r}
data <- read.csv("hw1_data.csv")
subs <- subset(data, Month == 6, select = Temp)
apply(subs, 2, mean)
```
</li>
</ol>

# Question 20
20. What was the maximum ozone value in the month of May (i.e. Month is equal to 5)?
<ol>
<li> **Answer:** 115 </li>
<li> **Explanation:**
```{r}
data <- read.csv("hw1_data.csv")
subs <- subset(data, Month == 5 & !is.na(Ozone), select = Ozone)
apply(subs, 2, max)
```
</li>
</ol>