Data Types
===================

Objects
-------

* character
* numeric (real numbers)
* integer
* complex
* logical (True / False)

The most basic object is a vector: can only contain lists of the same class
Empty vectors can be created using the vector() function

Numbers
-------

* Treated as numeric objects
* Integers are specified by adding "L" as a suffix

```{r}
x <- 1
class(x)
y <- 1L
class(y)
```

* There are special numbers, Inf representing infinity and NaN representing an undefined value (Not a Number)

```{r}
x <- 1/0
print(x)
y <- 0/0
print(y)
```

Attributes
----------

R objects can have attributes

* names, dimnames
* dimensions (e.g. matrices, arrays)
* class, length
* other user-defined attributes/metadata

Attributes of an object can be access using the attributes() function.

Creating Vectors
-----------------

The c() and vector() functions can be used to create vectors of objects.

```{r}
a <- c(0.5, 0.6)       # numeric
b <- c(TRUE, FALSE)    # logical
c <- c(T, F)           # logical
d <- c("a", "b", "c")  # character
e <- 9:29              # integer
f <- c(1+0i, 2+4i)     # complex

g <- vector("numeric", length=10)
print(g)
```

Mixing Objects
-------------------

```{r}
y <- c(1.7, "a") # character
y <- c(T, 0.6)   # numeric
y <- c("a", F)   # character
```

When different objects are mixed in a vector, coercion occurs so that every element in the vector is of the same class.

Explicit Coercion
-------------------

Objects can be explicitly coerced by using the as.* function

```{r}
x <- 0:6
class(x)
as.numeric(x)
as.logical(x)
as.character(x)
as.complex(x)
```

Nonsensical coercions result in NA

```{r warning=FALSE}
x <- c("a", "b", "c")
as.numeric(x)
as.logical(x)
as.complex(x)
```

Matrices
----------

Matrices are vectors with a dimension attribute. The dimension attribute is itself an integer vector of length 2 (nrow, ncol).

```{r}
m <- matrix(nrow = 2, ncol = 2); m
dim(m)
attributes(m)
```

Matrices are constructed column-wise, so entries can be thought of starting in the “upper left” corner and running down the columns.

```{r}
m <- matrix(1:6, nrow = 3, ncol = 2); m
```

Matrices can also be created directly from vectors by adding a dimension attribute.

```{r}
m <- 1:10; m
dim(m) <- c(2, 5); m
```

cbind-ing and rbind-ing
--------------------------

Matrices can be created by column-binding or row-binding with cbind() and rbind().

```{r}
x <- 1:3
y <- 10:12
cbind(x, y)
rbind(x, y)
```

Lists
--------

Lists are a special type of vector that can contain elements of different classes.

```{r}
x <- list("a", 3, T, 4L, 0+5i); x
x[5]
```

Factors
----------

Factors are used to represent categorical data. Factors can be unordered or ordered. One can think of a factor as an integer vector where each integer has a label.

```{r}
x <- factor(c("yes", "yes", "no", "yes", "no")); x
table(x)
unclass(x)
```

The order of the levels can be set using the levels argument to factor(). This can be important in linear modelling because the first level is used as the baseline level.

```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"),
            levels = c("yes", "maybe", "no"))
x
table(x)
unclass(x)
```

Missing Values
-----------------

Missing values are denoted by NA or NaN for undefined mathematical operations.

* is.na() is used to test objects if they are NA
* is.nan() is used to test for NaN
* NA values have a class also, so there are integer NA, character NA, etc.
* A NaN value is also NA but the converse is not true

```{r}
c <- c(1, 2, NA, 4, 5)
is.na(c)
is.nan(c)
d <- c(1, NaN, NA, 4, 5)
is.na(d)
is.nan(d)
```

Data Frames
--------------

Data frames are used to store tabular data

* They are represented as a special type of list where every element of the list has to have the same length
* Each element of the list can be thought of as a column and the length of each element of the list is the number of rows
* Unlike matrices, data frames can store different classes of objects in each column (just like lists); matrices must have every element be the same class
* Data frames also have a special attribute called row.names
* Data frames are usually created by calling read.table() or read.csv()
* Can be converted to a matrix by calling data.matrix()

```{r}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F)); x
nrow(x)
ncol(x)
```

Names Attribute
-----------------

R objects can also have names, which is very useful for writing readable code and self-describing objects.

```{r}
x <- 1:5; names(x)
names(x) <- c("Luna", "Yanna", "Samantha", "Via", "Kierra"); x
names(x)

# Lists can have names too!
x <- list(a = "d", b = "e", c = "f"); x

# So can matrices
x <- matrix(1:4, nrow=2, ncol=2)
dimnames(x) <- list(c("Amora", "Elyse"), c("Ledezma", "Camero")); x
```
